---
title: "Lab 10"
author: "Lili Xu"
date: "11/5/2021"
output: 
    html_document:
      html_preview: false
      keep_md: yes
    github_document:
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

## Setup

```{r set-up}
# install.packages(c("RSQLite", "DBI"))

library(RSQLite)
library(DBI)
library(dplyr)

# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
actor <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/actor.csv")
rental <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/rental.csv")
customer <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/customer.csv")
payment <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/payment_p2007_01.csv")

# Copy data.frames to database
dbWriteTable(con, "actor", actor)
dbWriteTable(con, "rental", rental)
dbWriteTable(con, "customer", customer)
dbWriteTable(con, "payment", payment)
```


Are the tables there?

```{r check-tables}
dbListTables(con)
```

You can also use knitr + SQL!

TIP: Use can use the following QUERY to see the structure of a table
```{sql, connection = con, output.var = "x1"}
PRAGMA table_info(actor)
```


This is equivalent to use `dbGetQuery`

#### Actor
```{r check-actor}
dbGetQuery(con, "PRAGMA table_info(actor)") %>%
  knitr::kable()
```

#### Customer
```{r check-customer}
dbGetQuery(con, "PRAGMA table_info(customer)")　%>%
  knitr::kable()
```

#### Rental
```{r check-rental}
dbGetQuery(con, "PRAGMA table_info(rental)") %>%
  knitr::kable()
```

#### Payment
```{r check-payment}
dbGetQuery(con, "PRAGMA table_info(payment)") %>%
  knitr::kable()
```

## Exercise 1
Retrive the actor ID, first name and last name for all actors using the actor table. Sort by last name and then by first name.

And using the LIMIT clause (`head()` in R ) to just show the first 5

```{r exercise1}
dbGetQuery(con, "
/* This is COMMENT! */
SELECT actor_id, first_name, last_name
FROM actor /* YOU CAN ADD COMMENTS USING
MULTIPLE LINES! */
ORDER by last_name, first_name 
LIMIT 5") %>%
  knitr::kable()

```

## Exercise 2
Retrive the actor ID, first name, and last name for actors whose last name equals ‘WILLIAMS’ or ‘DAVIS’.

```{r exercise2}

dbGetQuery(con, "
SELECT actor_id, first_name, last_name
FROM actor
WHERE last_name IN ('WILLIAMS', 'DAVIS')
LIMIT 5") %>%
  knitr::kable()
```


## Exercise 3
Write a query against the rental table that returns the IDs of the customers who rented a film on July 5, 2005 (use the rental.rental_date column, and you can use the date() function to ignore the time component). Include a single row for each distinct customer ID.


```{r exercise3}

dbGetQuery(con, "
SELECT DISTINCT customer_id
FROM rental
WHERE date(rental_date) = '2005-07-05'
LIMIT 5") %>%
  knitr::kable()

```


## Exercise 4


### Exercise 4.1
Construct a query that retrives all rows from the payment table where the amount is either 1.99, 7.99, 9.99.
```{r exercise4.1}

q <- dbSendQuery(con, "
SELECT *
FROM payment
WHERE amount IN (1.99, 7.99, 9.99)")

dbFetch(q , n = 10) %>%
  knitr::kable()

dbClearResult(q) 
```


### Exercise 4.2
Construct a query that retrives all rows from the `payment` table where the amount is greater then 5

```{r exercise4.2}
dbGetQuery(con, "
SELECT *
FROM payment
WHERE amount > 5
LIMIT 5") %>%
  knitr::kable()
```

#### Bouns: Count how many are

```{r bouns-count}
dbGetQuery(con, "
SELECT COUNT(*)
FROM payment
WHERE amount > 5") %>%
  knitr::kable()
```

Counting per `staff_id`

```{r Bouns-count-staff}
dbGetQuery(con, "
SELECT staff_id, COUNT(*) AS N
FROM payment
WHERE amount > 5
GROUP BY staff_id") %>%
  knitr::kable()
```

### Exercise 4.3
Construct a query that retrives all rows from the payment table where the amount is greater then 5 and less then 8

```{r exercise4.3}

dbGetQuery(con, "
SELECT *
FROM payment
WHERE amount > 5 AND amount < 8
LIMIT 5") %>%
  knitr::kable()

```


## Exercise 5
Retrive all the payment IDs and their amount from the customers whose last name is ‘DAVIS’.


```{r exercise5}
dbGetQuery(con, "
SELECT payment.payment_id, payment.amount
FROM payment
  INNER JOIN customer ON payment.customer_id = customer.customer_id
WHERE customer.last_name = 'DAVIS'") %>%
  knitr::kable()

```



## Exercise 6


### Exercise 6.1
Use COUNT(*) to count the number of rows in rental

```{r exercise6.1}
dbGetQuery(con, "
SELECT COUNT(*) as Total_Rentals
FROM rental") %>%
  knitr::kable()
```


### Exercise 6.2
Use COUNT(*) and GROUP BY to count the number of rentals for each customer_id

```{r exercise6.2}
dbGetQuery(con, "
SELECT customer_id, COUNT(*) AS 'N Rentals'
FROM rental
GROUP BY customer_id
LIMIT 10") %>%
  knitr::kable()
```


### Exercise 6.3
Repeat the previous query and sort by the count in descending order

```{r exercise6.3}
dbGetQuery(con, "
SELECT customer_id, COUNT(*) AS 'N Rentals'
FROM rental
GROUP BY customer_id
ORDER BY `N Rentals` DESC
LIMIT 10") %>%
  knitr::kable()
```



### Exercise 6.4
Repeat the previous query but use HAVING to only keep the groups with 40 or more.

```{r exercise6.4}
dbGetQuery(con, "
SELECT customer_id, COUNT(*) AS 'N Rentals'
FROM rental
GROUP BY customer_id
HAVING `N Rentals` >= 40
ORDER BY `N Rentals` DESC") %>%
  knitr::kable()
```


## Exercise 7

The following query calculates a number of summary statistics for the payment table using MAX, MIN, AVG and SUM

```{r exercise7}

dbGetQuery(con, "
SELECT 
  MAX(amount) AS Max, 
  MIN(amount) AS Min, 
  AVG(amount) AS Avg, 
  SUM(amount) AS Sum
FROM payment") %>%
  knitr::kable()

```


### Exercise 7.1
Modify the above query to do those calculations for each customer_id

```{r exercise7.1}
dbGetQuery(con, "
SELECT 
  customer_id, 
  MAX(amount) AS Max, 
  MIN(amount) AS Min,
  AVG(amount) AS Avg, 
  SUM(amount) AS Sum
FROM payment
GROUP BY customer_id
LIMIT 10") %>%
  knitr::kable()
```



### Exercise 7.2
Modify the above query to only keep the customer_ids that have more then 5 payments

```{r exercise7.2}
dbGetQuery(con, "
SELECT 
  customer_id, 
  MAX(amount) AS Max, 
  MIN(amount) AS Min,
  AVG(amount) AS Avg, 
  SUM(amount) AS Sum
FROM payment
GROUP BY customer_id
HAVING COUNT(customer_id) > 5") %>%
  knitr::kable()
```



## Clean up
Run the following chunk to disconnect from the connection.
```{r}
# clean up
dbDisconnect(con)
```

