---
title: "LAB3"
author: "Lili Xu"
date: "9/10/2021"
output: 
  html_document: 
    toc: yes
    toc_float: yes
    fig_width: 10
    fig_height: 10
    highlight: haddock
    theme: cosmo
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lab Description

The objective of the lab is to find the weather station with the highest elevation and look at patterns in the time series of its wind speed and temperature.

## 1. Read in the data


```{r download}
# install it only if it is not present
if(!require(data.table)){
  install.packages("data.table")
}


#only download it if file not exist
if (!file.exists("met_all.gz")){
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", 
                destfile = "met_all.gz", 
                method="libcurl", 
                timeout = 60)
}
met <- data.table::fread("met_all.gz")
```

## 2. Check the dimensions, headers, footers. 

How many columns, rows are there?


```{r check}
dim(met)
head(met)
tail(met)
```

There are 2,377,343 rows and 30 columns in the met dataset.

## 3. Take a look at the variables.

```{r structure}
str(met)
```

## 4. Take a closer look at the key variables.

```{r key}
table(met$year)
table(met$day)
table(met$hour)
summary(met$temp)
summary(met$elev)
summary(met$wind.sp)


### replace those miss data to NA
met[met$elev==9999.0] <- NA
summary(met$elev)

### remove the temperature below -40 which is another issue that contains minimum temperature
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)

```
Based on the information after clean out the data:
1. The weather station with highest elevation is 4113 meters.
2. There is a -17.2C temperature reading that seems suspicious.


## 5. Check the data against an external data source.

```{r checkAgainst}
met <- met[temp>-15]
met2 <- met[order(temp)]
head(met2)
```
Summarize that we removed temperatures colder than -15C. The new dataset has minimum temp -3C which is reasonable.


## 6. Calculate summary statistics

We want to pick out the weather station with maximum elevation and examine its windspeed and temperature.

```{r statistics}
elev <- met[elev==max(elev)]
summary(elev)

cor(elev$temp, elev$wind.sp, use="complete")

cor(elev$temp, elev$hour, use="complete")

cor(elev$wind.sp, elev$day, use="complete")

cor(elev$wind.sp, elev$hour, use="complete")

cor(elev$temp, elev$day, use="complete")
```


## 7. Exploratory graphs

```{r graphs, fig.height=9, fig.width=9}
hist(met$elev, breaks=100)
###### It seems elevation more around 0-500


hist(met$temp)
###### The temperature which is a normal Distribution, which is more focused on the middle, 20-26


hist(met$wind.sp)
###### The wind speed is more on low speed around 0-5


library(dplyr) ### use for function %>%
library(leaflet)### use for function leaflet
leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>%
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)
##### The location is based on the lat and lon



library(lubridate)
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)

elev <- elev[order(date)]
head(elev)
```


```{r fig.height=5, fig.width=10}
plot(elev$date, elev$temp, type='l')
```
The plot shows that the trend of the temperature is very stable, the frequency of the temperature difference in a month is close, and the daily temperature difference is relatively large, there is a difference of about 10 degrees. On August 11 and 12, the temperature difference of this month is the smallest and the coldest days.



```{r fig.height=5, fig.width=10}
plot(elev$date, elev$wind.sp, type='l')
```
According to the chart, the whole trend is very intense. It can be seen from the chart that the wind speed is still small from the beginning of the month to around August 10, and the difference is not large. Starting around August 11, the wind speed began to rise, and the difference in wind speed also increased, until around August 20, the wind speed began to decrease and became somewhat stable. Around August 24, the wind speed rose again and became intense, and the wind speed difference was even greater that day.
