---
title: "Lab6"
author: "Lili Xu"
date: "10/1/2021"
output: 
    html_document:
      toc: yes 
      toc_float: yes 
      highlight: haddock
      theme: cosmo
      keep_md: yes
    github_document:
      keep_html: false
      html_preview: false
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Download the Dataset

```{r pacakage}
library(tidyverse)
library(tibble)
library(dplyr)
library(ggplot2)
library(tidytext)
```


```{r get-data}
if (!file.exists("mtsamples.csv"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/00_mtsamples/mtsamples.csv",
    destfile = "mtsamples.csv",
    method = "libcurl",
    timeout = 60
  )
mtsamples <- read.csv("mtsamples.csv")
mtsamples <- as_tibble(mtsamples)
mtsamples
```




## Question 1: What specialties do we have?

We can use count() from dplyr to figure out how many different catagories do we have? Are these catagories related? overlapping? evenly distributed?

```{r dist-of-specialities}
specialties <- mtsamples %>%
  count(medical_specialty)

specialties %>%
  arrange(desc(n))%>%
  top_n(15) %>%
  knitr::kable()
  
```

There are `r nrow(speclialties)' specialties. Let's take a look at the distrubution.
```{r dist}
ggplot(mtsamples, aes(x = medical_specialty)) +
  geom_histogram(stat = "count") +
  coord_flip()

```
```{r dist2}
# method 2
ggplot(specialties, aes(x = n, y = fct_reorder(medical_specialty, n))) +
  geom_col()
```

There are not evenly ( uniformly) distribution.

## Question 2

```{r token-transcript,cache=TRUE}
mtsamples %>% 
  unnest_tokens(output = word, input = transcription) %>%
  count(word, sort = TRUE) %>%
  top_n(20) %>%
  ggplot(aes(x=n, y = fct_reorder(word,n))) +
    geom_col()
```

The word "patient" seems to be important, but we observe a lot of stopwords

## Question 3
Redo visualization but remove stopwords before
What do we see know that we have removed stop words? Does it give us a better idea of what the text is about?
```{r token-transcript-wo-stop,cache=TRUE}
mtsamples %>% 
  unnest_tokens(output = word, input = transcription) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words, by = c("word")) %>%
  top_n(20) %>%
  ggplot(aes(x=n, y = fct_reorder(word,n))) +
    geom_col()
```

Looking better ~~, but we don't like the numbers~~.

Bonus points if you remove numbers as well
```{r token-transcript-num,cache=TRUE}
# method 1
tokens_clean <- mtsamples %>% 
  unnest_tokens(output = word, input = transcription) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words, by = c("word"))

nums <- tokens_clean %>% filter(str_detect(word, "^[0-9]")) %>% select(word) %>% unique()

tokens_clean %>% 
  anti_join(nums, by = c("word")) %>%
  top_n(20) %>%
  ggplot(aes(x=n, y = fct_reorder(word,n))) +
    geom_col()

# method 2
mtsamples %>% 
  unnest_tokens(output = word, input = transcription) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words, by = c("word")) %>%
  # using regular expressions to remove numbers
  filter(!grepl(pattern = "^[0-9]+$", x = word)) %>%
  top_n(20) %>%
  ggplot(aes(x=n, y = fct_reorder(word,n))) +
    geom_col()

```



## Question 4

repeat question 2, but this time tokenize into bi-grams. how does the result change if you look at tri-grams?

```{r bi-grams-transcript}
mtsamples %>% 
  unnest_ngrams(output = bigram, input = transcription, n = 2) %>%
  count(bigram, sort = TRUE) %>%
  top_n(20) %>%
  ggplot(aes(x=n, y = fct_reorder(bigram,n))) +
    geom_col()
```

Using bi-grams is not very informative, Let's 

```{r tri-grams-transcript}
mtsamples %>% 
  unnest_ngrams(output = trigram, input = transcription, n = 3) %>%
  count(trigram, sort = TRUE) %>%
  top_n(20) %>%
  ggplot(aes(x=n, y = fct_reorder(trigram,n))) +
    geom_col()
```

Now some phrase start to show up, e.g., "tolerated the procedure", "prepped and draped".

## Question 5
```{r}

```

## Question 6
```{r}

```

## Question 7
```{r}

```
